\documentclass{article}
\usepackage{blindtext}
\usepackage{hyperref}
\title{State Constitution}
\date{May 2019}
\author{Iustitia Vincit}
\begin{document}
\maketitle
\newpage
\tableofcontents
\newpage
\section{Framework of Government}
\subsection{Deliberative Assembly}
\subsubsection{Formation}
\paragraph{Members}
The deliberative assembly shall consist of ninety-three local representatives and thirty-one party representatives.
\paragraph{Elections of Local Representatives}
Local representatives shall be elected every fourth year in their respective districts. They shall be divided into classes such that one-fourth is elected every second year.
\paragraph{Elections of Party Representatives}
Party representatives shall be elected every second year in a general election. Each political party or caucus shall be awarded a number of seats proportionate to their respective shares of the total vote.
\paragraph{Qualifications}
Each member of the deliberative assembly shall have been a citizen of the state for two years prior to the date of their elections. No person shall be elected without having attained eighteen years of age.
\paragraph{Majority Requirement and Quorum}
Except as otherwise provided by law, a two-thirds affirmative majority shall be necessary for the deliberative assembly to decide any question. A quorum shall consist of a requisite number of members for that majority, including each party representative or alternate. Absent members may be compelled to attend by penalties established by the rules of the deliberative assembly.
\subsubsection{State Governor and State Executive}
\paragraph{State Governor}
The state governor shall preside over the regular sessions of the deliberative assembly and see to its day-to-day business. The state governor shall be chosen by the local representatives in a secret election.
\paragraph{State Executive}
The state executive shall be nominated by the state governor and confirmed by the deliberative assembly.
\subparagraph{Succession of Office}
Upon the vacancy of the office or at other times during which the state executive is unable to perform the duties of office, the longest-serving member of the deliberative assembly shall assume the duties of office and undertake its powers until a successor is nominated and confirmed or the incumbent state executive regains the ability to execute the powers of office. During this time, the replacement state executive shall not exercise any powers or duties within the deliberative assembly.
\subparagraph{Relationship with State Councils}
The state executive shall serve as an \textit{ex officio} member of each state council. In this capacity, the state executive shall exercise no vote unless the members stand equally divided. The state executive shall convene and adjourn state councils as provided by law.
\subparagraph{General Powers}
The state executive may call forth the militia to execute the laws of the state, to suppress insurrection, and to act in the event of a natural disaster within the state. The state executive may grant pardons and reprieves for the crimes of the state with the advice and consent of the judicial council. The state executive may convene, preside over, and adjourn special sessions of the deliberative assembly and of the state councils.
\subparagraph{Enforcement Power}
The state executive shall enforce the laws of the state as the state executive councils provide.
\subsubsection{Legislative Process}
\paragraph{Legislative Committees}
For each pair of legislative and executive councils, there shall be a corresponding committee of the deliberative assembly.
\subparagraph{Committee Formation}
The members of each committee shall be drawn from the local representatives and assigned by the state governor. Each caucus of the deliberative assembly shall be assigned an equal number of seats on each committee. 
\subparagraph{Committee Considering of Bills}
Upon receiving a bill from its corresponding legislative council, each committee shall discuss and debate the bill. Amendments to the bill may be proposed and adopted by the committee during this time if approved by the supervisory judge of the corresponding legislative committee as complying with the purpose of the bill; decisions of the supervisory judge may be appealed to the court of appeals created for this purpose. The bill shall pass from the committee after ten days of receiving it unless the committee agrees to postpone, but cannot be postponed more than thirty days. The legislative process of the bill is terminated if the committee unanimously rejects the bill.
\paragraph{Assembly Considering of Bills}
Upon the passage of a bill from committee, the state governor shall schedule the bill for open discussion and debate for a time not less than seventy-two hours and in no less than three separate meetings from the deliberative assembly. Once the allotted discussion time has expired, the bill shall lay dormant for two days during which time representatives shall publicly issue any concerns and stand-asides or blocks along with the reasons therefor. After those two days, the state governor shall schedule a vote. If the deliberative assembly fails to approve the bill, it shall then vote to reject the bill. If both votes fail, the deliberative assembly shall be disbanded, all members recalled, and new elections held to chose successors to complete the terms of the previous members.
\paragraph{Stand-asides and Blocks}
Stand-asides shall not be considered in vote tallying, quorum, or majority requirements when voting on a bill. A block causes the immediate return of a bill to its originating committee for further amendment and later re-introduction, but if blocked again, the legislative process of that bill shall terminate. A block must be sustained by more than one-third of the local representatives and may be issued only by a party representative. Each party representative shall be entitled to one non-transferable and non-renewable block per each term, but may use that block twice on the same bill.
\subsubsection{Specific Powers}
\paragraph{Subpoenas}
The deliberative assembly or any committee may compel public officers, except judges, to provide sworn testimony.
\paragraph{Question Time}
The state governor shall hold question time once a week unless it is agreed to forego the event, during which the state governor and other officers of the deliberative assembly shall answer any questions of the members of the assembly.
\paragraph{Votes of No Confidence}
The deliberative assembly may dismiss either the state governor or the state executive upon a successful vote of no confidence. A motion of no confidence shall be brought only by a public petition bearing the signatures of one-tenth of the members of the deliberative assembly. Votes of no confidence shall not be scheduled within six months before an election of the deliberative assembly or less than four months after a previous unsuccessful vote of no confidence.
\paragraph{Impeachment}
The deliberative assembly may impeach any public officer of the state, including members of the assembly. Upon the adoption of an article of impeachment giving reasonable notice to the acts or omissions alleged to constitute impeachable offenses, a special state council shall be convened consisting of seven juridical nominees selected by lot. This council shall try the case against the impeached officer. The deliberative assembly shall elect two members to act as their representatives. The trial shall be conducted in the manner of civil proceedings and the impeached officer shall not be permitted to invoke protections against self-incrimination except as applicable in a general civil case. Conviction shall require the unanimous concurrence of the members of the council. Judgment shall not extend beyond removal from office and disqualification to hold any office of honor, trust, or profit of the state. The impeached officer, whether convicted or acquitted, shall be liable to prosecution and punishment pursuant to the law. The impeached officer shall not exercise the powers of office upon impeachment until acquittal.
\subsubsection{Snap Elections}
\paragraph{In General}
Snap elections of the deliberative assembly shall be held if a budget is not approved no less than two weeks prior to the expiration of the current budget, if the deliberative assembly is unable to elect a state governor or state executive for a period of time exceeding thirty days, or as otherwise provided by law.
\subsection{State Judiciary}
\subsubsection{State Supreme Court}
\paragraph{Formation and Vacancy}
The supreme court shall consist of thirty-five judges. Vacancies shall be filled by the state executive.
\paragraph{Trying Cases}
All cases from the lower courts shall first be certified to the supreme court and only later certified to the relevant court of appeals if the supreme court declines to try the case. The consent of one-third of judges shall be necessary to try case. Cases shall first be heard by a ten judge panel selected by lot. If this panel stands evenly divided or if a majority of the supreme court consents, cases may be appealed from this panel to an \textit{en banc} session of the supreme court.
\paragraph{Jurisdiction}
The supreme court shall have original jurisdiction in cases of \textit{quo warranto}, \textit{mandamus}, \textit{habeas corpus}, \textit{prohibition}, \textit{procedendo}, admission to the practice of law, and cases involving this constitution. The supreme court shall have general superintendent powers over all other courts of the state and may establish rules governing the practices of law and courts.
\subsubsection{State Courts of Appeals}
\paragraph{Formation and Vacancy}
Each court of appeals shall consist of three judges, except where the volume of business necessitates additional judges; cases shall be heard by a panel of three judges elected by lot. Vacancies shall be filled by the supreme court.
\paragraph{Jurisdiction}
Courts of appeals shall be constituted by the deliberative assembly with a given physical jurisdiction. Special courts of appeals may be established for certain types of cases where there is no trial court over which they shall have original jurisdiction. Courts of appeals shall have original jurisdiction in cases of \textit{quo warranto}, \textit{mandamus}, \textit{habeas corpus}, \textit{prohibition}, and \textit{procedendo}.
\paragraph{Appeal Conflicts}
Whenever the judges of a court of appeals find a judgment upon which they believe is in conflict with the judgment found by another court of appeals on the same question, they shall certify the case to the supreme court for final review and determination. 
\subsubsection{Lower Courts}
\paragraph{Trial Courts}
Trial courts shall be established by the deliberative assembly within the jurisdiction of each court of appeals. The deliberative assembly may devolve this power to other governing instruments.
\paragraph{Jurisdiction}
These courts shall have jurisdiction over all justiciable within their respective domains.
\subsubsection{Judicial Norms}
\paragraph{Reporting of Cases}
All cases and the reasons therefor shall be reported to the judicial council for official publication. All courts shall be courts of record.
\paragraph{Terms of Judges}
Judges shall serve life terms or until they no longer possess the qualifications of a juridical nominee.
\subsection{State Councils}
\subsubsection{Formation}
\paragraph{Establishment}
Each state council shall be established by the deliberative assembly with specific and particular competencies and powers. Unless otherwise provided, each council shall be comprised of the state executive, and one qualified person appointed by the state governor and each party representative who possess expertise in the competency of the council. 
\paragraph{Prohibition on Dual Office}
No person shall concurrently serve on two or more state councils.
\paragraph{Replacement of Members}
Members of a state council may be replaced by their respective appointors.
\subsubsection{Supervisory Judge}
\paragraph{Appointment}
Each state council shall be appointed a supervisory judge by the judicial council. Supervisory judges shall exercise no vote, but may assist in the drafting of policy.
\paragraph{Veto Power}
Supervisory judges may veto adopted policy on the basis of its intrinsic merits being in violation of this constitution. These vetoes may be appealed to the court of appeals created for this purpose by any member of the council.
\subsubsection{Constitutional Councils}
\paragraph{Legislative and Executive Councils}
The deliberative assembly may establish pairs of legislative and executive councils
\section{Framework of Rights}
\section{Framework of Law}
\end{document}